---
title: "TF depletion enrichment"
author: "Kushal K Dey"
date: "6/23/2017"
output: html_document
---

```{r}
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste((substring(s, 1,1)), tolower(substring(s, 2)),
      sep="", collapse=" ")
}
```


## mouse enrichment analysis

We check here whether depletion is enriched in TFs with specific functional annotations.

```{r}
all_mouse <- read.table("All Mouse names.txt")[,1]
depletion_mouse <- read.table("mouse.txt")[,1]
```

```{r}
all_mouse_names <- sapply(all_mouse, function(x) strsplit(as.character(x), "_")[[1]][1])
depletion_mouse_names <- sapply(depletion_mouse, function(x) strsplit(as.character(x), "_")[[1]][1])
```

```{r}
all_mouse_names_2 <- as.character(sapply(all_mouse_names, function(x) return(simpleCap(x))))
depletion_mouse_names_2 <- as.character(sapply(depletion_mouse_names, function(x) return(simpleCap(x))))
```

```{r}
write.table(all_mouse_names_2, file = "all_mouse_names.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(depletion_mouse_names_2, file = "depletion_mouse_names.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

No significant functional pathway was observed and no GO was found enriched at $0.01$ level of significance. 

## human enrichment analysis

We check here whether depletion is enriched in TFs with specific functional annotations.

```{r}
all_human <- read.table("All Human names.txt")[,1]
depletion_human <- read.table("Human.txt")[,1]
```

```{r}
all_human_names <- sapply(all_human, function(x) strsplit(as.character(x), "_")[[1]][1])
depletion_human_names <- sapply(depletion_human, function(x) strsplit(as.character(x), "_")[[1]][1])
```

```{r}
all_human_names_2 <- as.character(sapply(all_human_names, function(x) return(simpleCap(x))))
depletion_human_names_2 <- as.character(sapply(depletion_human_names, function(x) return(simpleCap(x))))
```

```{r}
write.table(all_human_names, file = "all_human_names.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(depletion_human_names, file = "depletion_human_names.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

```{r}
go <- read.delim2("go_human_depletion.tab")
```

No pathway was enriched but the following GO annotation was moderatley enriched (although the q value was well above $0.05$)

```{r}
go
```


