---
title: "Meeting0914"
author: "Dongyue Xie"
date: "September 13, 2017"
output: html_document
---

# Log scale of negative logo plot

Two types of heights. 1: IC based, split the IC based on the log scale norm. 2: Log based, the height of the logo is simply the log scale. 3. Log odds based

A comparison:
```{r,message=F,fig.width=20,fig.height=20}
library(Logolas)
mFile <- system.file("Exfiles/pwm1", package="seqLogo")
m <- read.table(mFile)
p <- seqLogo::makePWM(m)
pwm=p@pwm
color_profile=list("type" = "per_row","col" = RColorBrewer::brewer.pal(4,name ="Spectral"))

library(grid)
grid.newpage()
layout.rows <- 2
layout.cols <- 2
top.vp <- viewport(layout=grid.layout(layout.rows, layout.cols,
                                      widths=unit(rep(5,layout.cols), rep("null", 2)),
                                      heights=unit(rep(5,layout.rows), rep("null", 1))))
plot_reg <- vpList()
l <- 1
for(i in 1:layout.rows){
  for(j in 1:layout.cols){
    plot_reg[[l]] <- viewport(layout.pos.col = j, layout.pos.row = i, name = paste0("plotlogo", l))
    l <- l+1
  }
}


plot_tree <- vpTree(top.vp, plot_reg)

pushViewport(plot_tree)
seekViewport(paste0("plotlogo", 1))
nlogomaker(pwm,logoheight = 'ic',color_profile = color_profile,control = list(ic_scale=1),newpage = F)

seekViewport(paste0("plotlogo", 2))
nlogomaker(pwm,logoheight = 'log',color_profile = color_profile,
           pop_name='scale=1',control = list(depletion_weight=0,logscale=1),newpage = F)

seekViewport(paste0("plotlogo", 3))
nlogomaker(pwm,logoheight = 'log',
           pop_name='scale=0.1',color_profile = color_profile,control = list(depletion_weight=0,logscale=0.1),newpage = F)

seekViewport(paste0("plotlogo", 4))
nlogomaker(pwm,logoheight = 'log_odds',color_profile = color_profile,control = list(depletion_weight=0,log_odds_scale=1),newpage = F)
```

# Cancer mutation signature

[New plot](https://dongyuexie.github.io/Sequence-Motif-Depletion/mutation.html) - improved visualization.

Asked Yuchi - feedback

# Consensus seq

1. [Distance based](https://dongyuexie.github.io/Sequence-Motif-Depletion/consensus_seq.html)
2. Based on the [logo height](https://dongyuexie.github.io/Sequence-Motif-Depletion/nomenclature.html) of log scale. More intuitive when looking at the logo plot

Subjective

# Protein plot

Negative logo plot does not work well because many residues do not appear at one particular positions. Too many depletions.

PSSM plot, based on the value of PSSM. 

Protein database - [3FDB](http://caps.ncbs.res.in/3pfdb/index.html)
[Example](http://caps.ncbs.res.in/cgi-bin/mini/databases/3pfdb/get_entry.cgi?id=PF07388)

```{r,message=FALSE,warning=FALSE,echo=FALSE}
bg=c(0.074,0.052,0.045,0.054,0.025,0.034,0.054,0.074,0.026,0.068,0.099,0.058,0.025,0.047,0.039,0.057,0.051,0.013,0.034,0.073)
readprotein=function(file,skip=3,nrows,nsites,bg=bg,adash,mat){
  library(dash)
  rawdata=read.table(file = file,skip = skip,nrows = nrows)
  if(mat=='pwm'){
    pwm=as.matrix(rawdata[,23:42]/100)
  pfm=round(pwm*nsites)
  if(adash==T){
    pwm=dash(as.matrix(pfm),mode = bg,optmethod = 'mixEM')$posmean
  }
  colnames(pwm)=c('A' ,'R','N','D',   'C' ,  'Q',   'E' ,  'G',   'H' ,  'I',   'L' , 'K'  , 'M' ,  'F',   'P' ,  'S'  , 'T' ,  'W',   'Y',   'V')
  rownames(pwm)=1:nrow(pwm)
  return(t(pwm))
  }
  if(mat=='pssm'){
    pssm=as.matrix(rawdata[,3:22])
    colnames(pssm)=c('A' ,'R','N','D',   'C' ,  'Q',   'E' ,  'G',   'H' ,  'I',   'L' , 'K'  , 'M' ,  'F',   'P' ,  'S'  , 'T' ,  'W',   'Y',   'V')
  rownames(pssm)=1:nrow(pssm)
  return(t(pssm))
  }
}
pssm= readprotein('http://caps.ncbs.res.in/cgi-bin/mini/databases/3pfdb/3pfdb_pssm_download.cgi?id=PF07966&data_dir=SDB_folder',nrows = 29,nsites = 85,adash = FALSE,mat='pssm')
pwm= readprotein('http://caps.ncbs.res.in/cgi-bin/mini/databases/3pfdb/3pfdb_pssm_download.cgi?id=PF07966&data_dir=SDB_folder',nrows = 29,nsites = 85,adash = FALSE,mat='pwm')
pssm=pssm[,19:27]
colnames(pssm)=1:9
pwm=pwm[,19:27]
colnames(pwm)=1:9
```

```{r,message=FALSE,warning=FALSE}
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
RColorBrewer::brewer.pal(12, "Set3")[c(1,2,5,8,9)],
RColorBrewer::brewer.pal(9, "Set1")[c(9,7)],
RColorBrewer::brewer.pal(8, "Dark2")[c(3,4,8)])
color_profile <- list("type" = "per_row",
"col" = cols1)

nlogomaker(pwm,logoheight = 'ic',
color_profile = color_profile,
frame_width = 1)

logo_pssm(pssm,
          color_profile = color_profile,
          frame_width = 1
          )

```

# dash

1. [Intutition, model](https://dongyuexie.github.io/Sequence-Motif-Depletion/dashtutor.html)

2. example: small [sample size](https://dongyuexie.github.io/Sequence-Motif-Depletion/dashtutor.html), 
[combined dash](https://dongyuexie.github.io/Sequence-Motif-Depletion/app.html)

# vignette and website

[website](https://kkdey.github.io/Logolas-pages/)
