---
title: "TFBS Motif of plant"
author: "Dongyue Xie"
date: 2017-07-21
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

The database plantTFDB provides transcription factors of more than 160 species, including main lineages of green plants. What is special is that each species has specific background probability of A,C,G,T. Here, we firstly apply *dash* to the PFM and then plot the logo plot. Notice that only PWMs are provided so we first multiply the PWM with the nsites and the round to the nearest integer of each entry as the PFM. 
We compare the logo plots of the original PWM, the PWM from dash and the PWM from 'aggregated' dash. In the aggregated dash, we first combine all the PWMs of one species into a large matrix with four rows(of A,C,G,T) and $N_p$ columns where $N_p$ is the sum of the number of positions of each PWM.

We studied the TF binding motifs from 5 species with varying background probabilities, including  
Vigna radiata
Eucalyptus grandis
Arabidopsis thaliana
Actinidia chinensis.


```{r,message=FALSE,warning=FALSE}
library(Logolas)
# a function to get pwms, pfms, nsites etc from http://planttfdb.cbi.pku.edu.cn/
#'@ species: the name of speices, use abbreviation. Ach - Actinidia chinensis
#
plantTF=function(species){
  library(dash)
  #read all the TF names from the infomation page
  nam=readLines(paste('http://planttfdb.cbi.pku.edu.cn/download/motif/',species,'_TF_binding_motifs_information.txt',sep=''))
  
#record the original pwm from the website as a list
pwm_original=list()
#record the transformed pfm as a list
pfm_indi=list()
#record the pwm after applying the dash
pwm_dash=list()
#a large mattrix combine all the pfms
pfm_cb=c()
#record the number of positions for each pwm
npos=c()
#record the number of sequences of each pfm
site=c()

for(i in 2:length(nam)){
  #read each pwm file from the website 
  readfile=readLines(paste('http://planttfdb.cbi.pku.edu.cn/motif/Ach/',substring(nam[i],1,10),'.meme',sep = ''))
  #extract the pwm
  pwm=matrix(as.numeric(as.character(unlist(strsplit(readfile[13:(length(readfile)-3)],split = '\t')))),nrow = 4,byrow = F)
  rownames(pwm)=c('A','C','G','T')
  colnames(pwm)=1:ncol(pwm)
  pwm_original[[i-1]]=pwm
  #extract the number of sites
  nsites=as.numeric(as.character(unlist(strsplit(readfile[12],split = ' '))[grep('nsites',unlist(strsplit(readfile[12],split = ' ')))+1]))
  site[i-1]=nsites
  #extract the background probabilities
  bg=as.numeric(strsplit(readfile[8],split = ' ')[[1]][c(2,4,6,8)])
  #
  pfm=round(pwm*nsites)
  pfm_cb=cbind(pfm_cb,pfm)
  npos[i-1]=ncol(pwm)
  pfm_indi[[i-1]]=pfm
  pwm=dash(t(pfm),optmethod = 'mixEM',mode = bg)
  pwm=t(pwm$posmean)
  rownames(pwm)=c('A','C','G','T')
  colnames(pwm)=1:ncol(pwm)
  pwm_dash[[i-1]]=pwm
}
pwm_adash=dash(t(pfm_cb),mode=bg,optmethod = 'mixEM')
pwm_adash=t(pwm_adash$posmean)
pwm_agg=list()
for(i in 1:length(npos)){
  pwm=pwm_adash[,1:npos[i]]
  rownames(pwm)=c('A','C','G','T')
  colnames(pwm)=1:ncol(pwm)
  pwm_adash=pwm_adash[,-(1:npos[i])]
  pwm_agg[[i]]=pwm
}

results=list(pfm=pfm_indi,pwm_original=pwm_original,pwm_dash=pwm_dash,pwm_cbdash=pwm_agg,bg=bg,nseq=site,npos=npos)

return(results)

}

```

Then we show the logo plots and negative logo plots for the three PWMs from 
Actinidia chinensis. We pick the PWMs whose the number of sequences are small.

```{r,message=FALSE,warning=FALSE}

Ach=plantTF('Ach')

color_profile = list("type" = "per_row", 
                     "col" = RColorBrewer::brewer.pal(4,name ="Spectral"))
set.seed(1)
for (i in sample(which(Ach$nseq<=20),5)){
  
   logomaker(Ach$pwm_original[[i]],color_profile = color_profile,frame_width = 1,bg=Ach$bg,pop_name = paste('original pwm',i))
  logomaker(Ach$pwm_dash[[i]],color_profile = color_profile,frame_width = 1,bg=Ach$bg,pop_name = paste('dash pwm',i))
  logomaker(Ach$pwm_cbdash[[i]],color_profile = color_profile,frame_width = 1,bg=Ach$bg,pop_name = paste('combined dash pwm',i))
  
nlogomaker(Ach$pwm_original[[i]],color_profile = color_profile,frame_width = 1,logoheight = 'log_odds',bg=Ach$bg,control = list(logscale = 0.2, quant = 0.5,depletion_weight = 0.5),pop_name = paste('original pwm',i))
  nlogomaker(Ach$pwm_dash[[i]],color_profile = color_profile,frame_width = 1,logoheight = 'log_odds',bg=Ach$bg,control = list(logscale = 0.2, quant = 0.5,depletion_weight = 0.5),pop_name = paste('dash pwm',i))
  nlogomaker(Ach$pwm_cbdash[[i]],color_profile = color_profile,frame_width = 1,logoheight = 'log_odds',bg=Ach$bg,control = list(logscale = 0.2, quant = 0.5,depletion_weight = 0.5),pop_name = paste('combined dash pwm',i))
 
}

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
