---
title: "Untitled"
author: "Dongyue Xie"
date: "September 9, 2017"
output: html_document
---

# Creating logos

Logolas provides users the freedom to create specific logos. Take the symbol "Lambda" for example. Firstly, create the function for it:

```{r,warning=FALSE,message=FALSE}
library(Logolas)
LAMBDAletter <- function(plot=FALSE,
                         fill_symbol = TRUE,
                         colfill="green",
                         lwd=10){

  x <- c(0.15, 0.5, 0.85, 0.75, 0.5, 0.25)
  y <- c(0, 1, 0, 0, 0.8, 0)
  fill <- colfill
  id <- rep(1, length(x))

  colfill <- rep(colfill, length(unique(id)))

  if(plot){
    get_plot(x, y, id, fill, colfill, lwd = lwd, fill_symbol = fill_symbol)
  }

  ll <- list("x"= x,
             "y"= y,
             "id" = id,
             "fill" = fill,
             "colfill" = colfill)
  return(ll)
}

```

Notice that the function name has to be of the form "$*$letter", where the user can be creative with the ”$*$” part. Also the name must be in uppercase letters. The user can then check if the symbol plot looks like a lambda or not in the following way.


```{r,warning=FALSE,message=FALSE,fig.width=4,fig.height=2}

lambda <- LAMBDAletter()
grid::grid.newpage()
grid::pushViewport(grid::viewport(x=0.5,y=0.5,width=1, height=1,
                                  clip=TRUE))
grid::grid.polygon(lambda$x, lambda$y,
                   default.unit="native",
                   id=lambda$id,
                   gp=grid::gpar(fill=lambda$fill,
                                 lwd=10))


```

The user can then add this symbol to the library. To use lambda as part of a string, the user has to put lambda inside ”/.../” to make sure that the function reads it as a new symbol and not general English alphabets or numbers. We provide an example below.

